---
title: "PA1_template"
author: "Hamid R. Darabi"
date: "Tuesday, August 12, 2014"
output: html_document
---

In the first part of this assignment I want to import the data.

```{r}
library(plyr)
unzip("activity.zip")
df  <- read.csv("activity.csv")
dfn <- df[ complete.cases(df), ]
ddata <- ddply(dfn, .(date) , summarize , mean = mean(steps) )
hist(ddata$mean, breaks=10, col="green", main="Average Daily Steps Histogram", xlab="Number of Steps")
ddata <- ddply(dfn, .(date) , summarize , sum = sum(steps) )
sprintf("Mean of daily steps is %0.2f and median is %.02f", mean(ddata$sum), median(ddata$sum), digits=2)
oldmean <- mean(ddata$sum)
oldmedian <- median(ddata$sum)
```
In the second section of the assignment the daily pattern of steps is analyzed.
```{r}
ddata <- ddply(dfn, .(interval) , summarize , mean = mean(steps) )
plot.ts( ddata$mean , main = "Average Steps per Time Interval")
sprintf("Maximum number of steps is on : %d", ddata[ ddata$mean == max(ddata$mean) ,]$interval )
```
In this section I need to deal with missing values. I replace them with the same interval mean from all days.
```{r}
sprintf("There are %d rows with missing values.", nrow(df)-nrow(dfn))
ddata <- ddply(dfn, .(interval) , summarize , mean = mean(steps) )
dfnomis <- merge(df, ddata)
dfnomis[ is.na(dfnomis$steps),"steps"] <- dfnomis[is.na(dfnomis$steps),"mean"] 
dfnomis$mean <- NULL
ddata <- ddply(dfn, .(date) , summarize , sum = sum(steps) )
hist(ddata$sum, breaks=10, col="green", main="Average Daily Steps Histogram", xlab="Number of Steps")
sprintf("Mean of daily steps is %0.2f and median is %.02f", mean(ddata$sum), median(ddata$sum), digits=2)
sprintf("The changes because of filling missing values are: mean: %f and median %f", mean(ddata$sum)-oldmean, median(ddata$sum)-oldmedian )
```
The last section of the assignment is about the pattern change from weekdays to weekends. I plot two timeseries representing each one to compare them.
```{r}
dfnomis$date <- as.Date(dfnomis$date)
dfnomis$isweekend <- factor( weekdays(dfnomis$date) %in% c("Sunday", "Saturday"), labels=c("weekday","weekend"))
selected <- dfnomis[dfnomis$isweekend == "weekday" ,]
ddata <- ddply(selected, .(interval) , summarize , mean = mean(steps) )
par(mfrow=c(2,1),mar = rep(2, 4))
plot.ts( ddata$mean , main = "Average Steps in Weekdays")
selected <- dfnomis[dfnomis$isweekend == "weekend" ,]
ddata <- ddply(selected, .(interval) , summarize , mean = mean(steps) )
plot.ts( ddata$mean , main = "Average Steps in Weekends")
```
